---
title: "Monitoring Americans' Mental Health under Lockdown (Updated Daily)"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(plotly)
library(tidyr)
library(leaflet)
library(gtrendsR)
library(magrittr)
library(tidyverse)
library(rsconnect)

##### Google Trends

### Locked down states

lock_states <- list("US-CA", "US-IL", "US-CT", "US-NY", "US-NJ", "US-DE", "US-OH", "US-LA", "US-NV", "US-FL", "US-WV", "US-RI", "US-PA", "US-OR", "US-MI", "US-MA", "US-MD",  "US-IN", "US-CO", "US-HI", "US-ID", "US-MN", "US-NM", "US-OK", "US-TX", "US-VA", "US-WA", "US-WI")
lock_names <- list("ca_time", "il_time", "ct_time", "ny_time", "nj_time", "de_time", "oh_time", "la_time", "nv_time", "fl_time", "ind_time", "md_time", "ma_time", "mi_time", "or_time", "pa_time", "ri_time", "wv_time", "co_time", "hi_time", "id_time", "mn_time", "nm_time", "ok_time", "tx_time", "va_time", "wa_time", "wi_time")

### Not locked down states

not_lock_states <- list("US-AL", "US-AK", "US-AZ", "US-AR", "US-GA", "US-IA", "US-KS", "US-KY", "US-ME", "US-MS", "US-MO", "US-MT", "US-NE", "US-NH", "US-NC", "US-ND", "US-SC", "US-SD", "US-TN", "US-UT", "US-VT", "US-WY")
not_lock_names <- list("al_time", "ak_time", "az_time", "ar_time", "ga_time", "ia_time", "ks_time", "ky_time", "me_time", "ms_time", "mo_time", "mt_time", "ne_time", "nh_time", "nc_time", "nd_time", "sc_time", "sd_time", "tn_time", "ut_time", "vt_time", "wy_time")


##### Map 
### Cases

confirm <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv")

confirm <- confirm[which(confirm$Country.Region=="US"), ]
confm <- confirm[1:52, ]
confm <- confm[-c(5, 6), ]
confm <- confm[, c(1:4, length(confm))]

pal <- colorFactor(c("#E95555"), domain = c("California", "New York", "Connecticut", "New Jersey", "Delaware", "Illinois", "Florida", "Nevada", "Louisiana", "Ohio", "West Virginia", "Rhode Island", "Pennsylvania", "Oregon", "Michigan", "Massachusetts", "Maryland", "Indiana"), na.color = "#81C77F")

### Deaths (not used here for now)

#deaths_all <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Deaths.csv")
#deaths <- deaths_all[which(deaths_all$Country.Region=="US"), ]
#deaths <- deaths[1:52, ]
#deaths <- deaths[-c(5, 6), ]
#deaths <- deaths[, c(1:4, length(deaths))]

```

Graphs 
=====================================     

Row {data-height=500}
-----------------------------------------------------------------------

### Search related to "Depression"

```{r}

### Locked down

l_list <- list()
for (i in lock_states) {
  ls <- gtrendsR::gtrends(keyword="Depression", geo=i, time="today 1-m")
  ls <- data.frame(ls$interest_over_time["date"], ls$interest_over_time["hits"])
  for (n in lock_names) {
    n <- ls
    l_list[[i]] <- n
  }
}

lock_down <- Reduce(function(x, y) merge(x, y, all=T,by.x="date", by.y="date"), l_list, accumulate=F)
lock_down$average <- apply(lock_down[, 2:ncol(lock_down)], 1, mean)
lock_down_final <- data.frame(lock_down$date, lock_down$average)
colnames(lock_down_final) <- c("dates", "lock_average")

### Not locked down

n_list <- list()
for (p in not_lock_states) {
  nls <- gtrendsR::gtrends(keyword="Depression", geo=p, time="today 1-m")
  nls <- data.frame(nls$interest_over_time["date"], nls$interest_over_time["hits"])
  for (k in not_lock_names) {
    k <- nls
    n_list[[p]] <- k
  }
}

not_lock_down <- Reduce(function(x, y) merge(x, y, all=T,by.x="date", by.y="date"), n_list, accumulate=F)
not_lock_down$average <- apply(not_lock_down[, 2:ncol(not_lock_down)], 1, mean)
not_lock_down_final <- data.frame(not_lock_down$date, not_lock_down$average)
colnames(not_lock_down_final) <- c("dates", "lock_average")

### Merge lock down and not lock down data

both <- merge(lock_down_final, not_lock_down_final, by="dates")
colnames(both) <- c("dates", "lock_average", "not_lock_average")

both_data <- both %>%
  select(dates, lock_average, not_lock_average) %>%
  gather(key = "types", value = "interest", -dates)

### Graph

lock_graph <- ggplot(both_data, aes(x=dates, y=interest)) + 
  geom_line(aes(linetype=types, size=types, color=types)) + 
  labs(x="Dates", y='"Depression" (average)') +
  scale_color_manual(values=c("#E95555", "#81C77F"), name='U.S. states under: ', labels=c("Lock down", "Not lock down")) +
  scale_size_manual(values=c(1.2, 1.2), name='U.S. states under: ', labels=c("Lock down", "Not lock down")) +
  scale_linetype_manual(values=c("solid", "solid"), name='U.S. states under: ', labels=c("Lock down", "Not lock down")) +
  theme(legend.position="top", legend.title=element_text(size=12), legend.text=element_text(size=12), axis.title.y=element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)), axis.title.x=element_text(margin=margin(t=11, r=0, b=0, l=0))) 

lock_graph


```

### U.S. Coronavirus Cases by State

```{r}

### Map

pal2 <- colorFactor(
  palette = c("#E95555", "#81C77F"),
  domain = c("Lock down", "Not lock down")
)

leaflet(confm) %>% addTiles() %>%
  addCircleMarkers(lng = ~Long, lat = ~Lat, weight = 1,
    radius = ~sqrt(confm[, length(confm)])/1.7, popup = paste("State: ", confm$Province.State, "<br>", "Cases: ", confm[, 5]), stroke=F, fillColor = ~pal(Province.State), fillOpacity = 0.5
  ) %>%
  addLegend( pal = pal2, values = c("Lock down", "Not lock down"),
    labFormat = labelFormat(prefix = ""),
    opacity = 1
  ) %>%
  setView(lat = 37.0902, lng = -95.7129, zoom = 3)




```

Row {data-height=500}
-----------------------------------------------------------------------

### Search related to "Domestic Violence"

```{r}

### Locked down

l_list <- list()
for (i in lock_states) {
  ls <- gtrendsR::gtrends(keyword="Domestic violence", geo=i, time="today 1-m")
  ls <- data.frame(ls$interest_over_time["date"], ls$interest_over_time["hits"])
  for (n in lock_names) {
    n <- ls
    l_list[[i]] <- n
  }
}

lock_down <- Reduce(function(x, y) merge(x, y, all=T,by.x="date", by.y="date"), l_list, accumulate=F)
lock_down$average <- apply(lock_down[, 2:ncol(lock_down)], 1, mean)
lock_down_final <- data.frame(lock_down$date, lock_down$average)
colnames(lock_down_final) <- c("dates", "lock_average")

### Not locked down

n_list <- list()
for (p in not_lock_states) {
  nls <- gtrendsR::gtrends(keyword="Domestic violence", geo=p, time="today 1-m")
  nls <- data.frame(nls$interest_over_time["date"], nls$interest_over_time["hits"])
  for (k in not_lock_names) {
    k <- nls
    n_list[[p]] <- k
  }
}

not_lock_down <- Reduce(function(x, y) merge(x, y, all=T,by.x="date", by.y="date"), n_list, accumulate=F)
not_lock_down$average <- apply(not_lock_down[, 2:ncol(not_lock_down)], 1, mean)
not_lock_down_final <- data.frame(not_lock_down$date, not_lock_down$average)
colnames(not_lock_down_final) <- c("dates", "lock_average")

### Merge lock down and not lock down data

both <- merge(lock_down_final, not_lock_down_final, by="dates")
colnames(both) <- c("dates", "lock_average", "not_lock_average")

both_data <- both %>%
  select(dates, lock_average, not_lock_average) %>%
  gather(key = "types", value = "interest", -dates)

### Graph

lock_graph <- ggplot(both_data, aes(x=dates, y=interest)) + 
  geom_line(aes(linetype=types, size=types, color=types)) + 
  labs(x="Dates", y='"Domestic violence" (average)') +
  scale_color_manual(values=c("#E95555", "#81C77F"), name='U.S. states under: ', labels=c("Lock down", "Not lock down")) +
  scale_size_manual(values=c(1.2, 1.2), name='U.S. states under: ', labels=c("Lock down", "Not lock down")) +
  scale_linetype_manual(values=c("solid", "solid"), name='U.S. states under: ', labels=c("Lock down", "Not lock down")) +
  theme(legend.position="top", legend.title=element_text(size=12), legend.text=element_text(size=12), axis.title.y=element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)), axis.title.x=element_text(margin=margin(t=11, r=0, b=0, l=0))) 

lock_graph

```


### Search related to "Suicide"

```{r}

### Locked down

l_list <- list()
for (i in lock_states) {
  ls <- gtrendsR::gtrends(keyword="Suicide", geo=i, time="today 1-m")
  ls <- data.frame(ls$interest_over_time["date"], ls$interest_over_time["hits"])
  for (n in lock_names) {
    n <- ls
    l_list[[i]] <- n
  }
}

lock_down <- Reduce(function(x, y) merge(x, y, all=T,by.x="date", by.y="date"), l_list, accumulate=F)
lock_down$average <- apply(lock_down[, 2:ncol(lock_down)], 1, mean)
lock_down_final <- data.frame(lock_down$date, lock_down$average)
colnames(lock_down_final) <- c("dates", "lock_average")

### Not locked down

n_list <- list()
for (p in not_lock_states) {
  nls <- gtrendsR::gtrends(keyword="Suicide", geo=p, time="today 1-m")
  nls <- data.frame(nls$interest_over_time["date"], nls$interest_over_time["hits"])
  for (k in not_lock_names) {
    k <- nls
    n_list[[p]] <- k
  }
}

not_lock_down <- Reduce(function(x, y) merge(x, y, all=T,by.x="date", by.y="date"), n_list, accumulate=F)
not_lock_down$average <- apply(not_lock_down[, 2:ncol(not_lock_down)], 1, mean)
not_lock_down_final <- data.frame(not_lock_down$date, not_lock_down$average)
colnames(not_lock_down_final) <- c("dates", "lock_average")

### Merge lock down and not lock down data

both <- merge(lock_down_final, not_lock_down_final, by="dates")
colnames(both) <- c("dates", "lock_average", "not_lock_average")

both_data <- both %>%
  select(dates, lock_average, not_lock_average) %>%
  gather(key = "types", value = "interest", -dates)

### Graph

lock_graph <- ggplot(both_data, aes(x=dates, y=interest)) + 
  geom_line(aes(linetype=types, size=types, color=types)) + 
  labs(x="Dates", y='"Suicide" (average)') +
  scale_color_manual(values=c("#E95555", "#81C77F"), name='U.S. states under: ', labels=c("Lock down", "Not lock down")) +
  scale_size_manual(values=c(1.2, 1.2), name='U.S. states under: ', labels=c("Lock down", "Not lock down")) +
  scale_linetype_manual(values=c("solid", "solid"), name='U.S. states under: ', labels=c("Lock down", "Not lock down")) +
  theme(legend.position="top", legend.title=element_text(size=12), legend.text=element_text(size=12), axis.title.y=element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)), axis.title.x=element_text(margin=margin(t=11, r=0, b=0, l=0))) 

lock_graph

```

About
===================================== 

**Purpose**

More U.S. states are entering lockdown. Under lockdown, people can't go outside except for "essential" businesses. These businesses include pharmacy, hospital, and grocery that provide sustenance. They exclude dine-in restaurants, parks, and gyms where people usually gather to socialize. The graphs show how often people in each type of states -- locked down or not locked down -- search terms that represent change in their mental health and living environment. The map gives context. 

How the graphs change over time may reveal the cost of lockdown other than the economic: the mental cost. 

This isn't to advocate against the lockdown but to encourage finding solutions to the virus and to the mental cost.

**Data sources**

1. 2019 Novel Coronavirus COVID-19 (2019-nCoV) Data Repository by Johns Hopkins CSSE (https://github.com/CSSEGISandData/COVID-19)

2. Google Trend (https://trends.google.com/trends/?geo=US)

All data used here are updated daily. Each graph shows the data for the past 30 days from today. The map shows all coronavirus cases in the U.S. for each state. Clicking on a circle shows the name of the state and the total number of its cases. 

**States under lockdown** 

* California
* New York
* Connecticut
* New Jersey
* Delaware
* Illinois
* Florida
* Nevada
* Louisiana
* Ohio
* West Virginia
* Rhode Island
* Pennsylvania
* Oregon
* Michigan
* Massachusetts
* Maryland
* Indiana
  
**Contact**

For any inquiries/feedback, please email David: slee511@jhu.edu. 