---
title: "Do Masks Help?"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    runtime: shiny
---
Graphs 
=====================================     

```{r setup, include=FALSE}
library(flexdashboard)
library(ggplot2)
library(plotly)
library(tidyr)
library(leaflet)
library(magrittr)
library(tidyverse) 
library(rsconnect) 
library(shiny)

cases <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")

### Countries with masks

china <- cases[which(cases$Country.Region=="China"), ]
china <- within(china, rm("Province.State", "Lat", "Long"))
china <- colSums(china[, 2:length(china)])
korea <- cases[which(cases$Country.Region=="Korea, South"), ]
korea <- korea[, 5:length(korea)]
singapore <- cases[which( cases$Country.Region=="Singapore"), ]
singapore <- singapore[, 5:length(singapore)]

### Countries advocating against masks

us <- cases[which(cases$Country.Region=="US"), ]
us <- us[, 5:length(us)]
italy <- cases[which(cases$Country.Region=="Italy"), ]
italy <- italy[, 5:length(italy)]
spain <- cases[which(cases$Country.Region=="Spain"), ]
spain <- spain[, 5:length(spain)]
france <- cases[117, ]
france <- france[, 5:length(france)]
britain <- cases[224, ]
britain <- britain[, 5:length(britain)]

cases <- rbind(us, italy, spain, france, britain, china, korea, singapore)
cases <- t(cases)
cases <- data.frame(row.names(cases), cases, row.names=NULL)
names(cases) <- c("Time", "US", "Italy", "Spain", "France", "UK", "China", "South.Korea", "Singapore")

#cases$Countries <- c("US", "Italy", "Spain", "China", "South Korea", "Singapore")

cases$Dates <- as.Date(strptime(cases$Time, 'X%m.%d'))


### Deaths

casesd <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")

chinad <- casesd[which(casesd$Country.Region=="China"), ]
chinad <- within(chinad, rm("Province.State", "Lat", "Long"))
chinad <- colSums(chinad[, 2:length(chinad)])
koread <- casesd[which(casesd$Country.Region=="Korea, South"), ]
koread <- koread[, 5:length(koread)]
singapored <- casesd[which(casesd$Country.Region=="Singapore"), ]
singapored <- singapored[, 5:length(singapored)]

### Countries advocating against masks

usd <- casesd[which(casesd$Country.Region=="US"), ]
usd <- usd[, 5:length(usd)]
italyd <- casesd[which(casesd$Country.Region=="Italy"), ]
italyd <- italyd[, 5:length(italyd)]
spaind <- casesd[which(casesd$Country.Region=="Spain"), ]
spaind <- spaind[, 5:length(spaind)]
franced <- casesd[117, ]
franced <- franced[, 5:length(franced)]
britaind <- casesd[224, ]
britaind <- britaind[, 5:length(britaind)]

casesd <- rbind(usd, italyd, spaind, franced, britaind, chinad, koread, singapored)
casesd <- t(casesd)
casesd <- data.frame(row.names(casesd), casesd, row.names=NULL)
names(casesd) <- c("Time", "US", "Italy", "Spain", "France", "UK", "China", "South.Korea", "Singapore")

casesd$Dates <- as.Date(strptime(casesd$Time, 'X%m.%d'))


```

Col {data-width=500}
-----------------------------------------------------------------------

### 

```{r}

#m = list( l = 80, r = 300, b = 80, t = 100, pad = 0 )

cg <- ggplot(cases) +
  geom_path(aes(x=Dates, y=US, group=1, linetype="Limited masks"), color="coral2") +
  geom_path(aes(x=Dates, y=Italy, group=1, linetype="Limited masks"), color="coral2") +
  geom_path(aes(x=Dates, y=Spain, group=1, linetype="Limited masks"), color="coral2") +
  geom_path(aes(x=Dates, y=France, group=1, linetype="Limited masks"), color="coral2") + 
  geom_path(aes(x=Dates, y=UK, group=1, linetype="Limited masks"), color="coral2") + 
  geom_path(aes(x=Dates, y=China, group=1, linetype="Many masks"), color="lightcyan4") +
  geom_path(aes(x=Dates, y=South.Korea, group=1, linetype="Many masks"), color="lightcyan4") +
  geom_path(aes(x=Dates, y=Singapore, group=1, linetype="Many masks"), color="lightcyan4") + 
  ggtitle("COVID-19 Cases by Country") +
  scale_linetype_manual(values=c("#E95555", "#81C77F")) +
  labs(x="Time", y='COVID-19 Cases by Country')  +
  theme(plot.title = element_text(color="black", size=12, face="", hjust = 0.5), legend.position=c(.1, .8), legend.text=element_text(size=10, face="Avenir"), axis.title.y=element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)), axis.title.x=element_text(margin=margin(t=11, r=0, b=0, l=0))) 
  #theme(legend.title=element_blank()) 
cg <- ggplotly(cg) %>% layout(autosize=T, legend = list(x =0.01, y = .99), xaxis=element_text(margin=margin(t=11, r=0, b=0, l=0)))
cg





```

### 

* Singapore, China, and South Korea have flooded their countries with masks. The US, Italy, France, UK, and Spain have not. The graphs show the difference (updated daily).

* Hover over the lines to see the details.

* Wall Street Journal on how the virus spreads: "Thousands of tiny, often invisible droplets of saliva or mucus disperse in the air. You walk by -- within 6 feet of the offender -- and inadvertently inhale the droplets." (https://www.wsj.com/articles/how-the-coronavirus-attacks-your-body-11585343549)

* George Gao, director-general of the Chinese Center for Disease Control and Prevention, who battled early COVID-19 spread in China: "The big mistake in the U.S. and Europe, in my opinion, is that people aren’t wearing masks. This virus is transmitted by droplets and close contact. Droplets play a very important role—you’ve got to wear a mask..." (https://www.sciencemag.org/news/2020/03/not-wearing-masks-protect-against-coronavirus-big-mistake-top-chinese-scientist-says)


Col {data-width=500}
-----------------------------------------------------------------------

### 

```{r}

cg <- ggplot(casesd) +
  geom_path(aes(x=Dates, y=US, group=1, linetype="Limited masks"), color="coral2") +
  geom_path(aes(x=Dates, y=Italy, group=1, linetype="Limited masks"), color="coral2") +
  geom_path(aes(x=Dates, y=Spain, group=1, linetype="Limited masks"), color="coral2") +
  geom_path(aes(x=Dates, y=France, group=1, linetype="Limited masks"), color="coral2") + 
  geom_path(aes(x=Dates, y=UK, group=1, linetype="Limited masks"), color="coral2") + 
  geom_path(aes(x=Dates, y=China, group=1, linetype="Many masks"), color="lightcyan4") +
  geom_path(aes(x=Dates, y=South.Korea, group=1, linetype="Many masks"), color="lightcyan4") +
  geom_path(aes(x=Dates, y=Singapore, group=1, linetype="Many masks"), color="lightcyan4") + 
  ggtitle("Deaths from COVID-19 by Country") +
  scale_linetype_manual(name="Deaths", values=c("#E95555", "#81C77F")) +
  labs(x="Time", y='COVID-19 Deaths by Country')  +
  theme(plot.title = element_text(color="black", size=12, face="", hjust = 0.5), legend.position=c(.1, .8), legend.text=element_text(size=10, face="Avenir"), axis.title.y=element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)), axis.title.x=element_text(margin=margin(t=11, r=0, b=0, l=0))) 
  #theme(legend.title=element_text("Hello")) 
cg <- ggplotly(cg) %>% layout(autosize=T, legend = list(x =0.01, y = .99), xaxis=element_text(margin=margin(t=11, r=0, b=0, l=0)))
cg

```

### 

* Americans and Europeans have been diligently washing their hands. COVID-19 has still ravaged the countries. 

* In the many-masks countries, people are publicly shamed if they don't wear masks. In the limited-masks countries, if they wear masks (perceived as sick).

* The disparity between the two types in rate of cases and of deaths is undeniable.

* Masks also have indirect impact; wearing them tells others to take the crisis seriously and avoid contact. 

* Masks should be worn not only on the streets but also in buildings, elevators, etc. If people are there, droplets also are. 

* But the shortage is real. I believe homemade masks are the answer. Today (March 31) NYT columnist Farhad Manjoo wrote this about homemade masks: https://www.nytimes.com/2020/03/31/opinion/coronavirus-n95-mask.html. They (and simple scarves) are incredibly effective. Better to donate manufactured masks to healthcare workers, who need them to reassure patients. 

* (Update, March 31) CDC, WH, etc. are revisiting their recommendation of not wearing masks in public (I sent WH and others this website on March 29). 


Contact
=====================================     

**Data sources**

* 2019 Novel Coronavirus COVID-19 (2019-nCoV) Data Repository by Johns Hopkins CSSE (https://github.com/CSSEGISandData/COVID-19)

The data used here are updated daily. 

**Contact**

For any inquiries, please email David Lee: slee511@jhu.edu
